### 2013-06-26 Tex/chapters/monte_carlo.tex
weitere Anmerkungen zu verschiedenen MC algorithmen

\subsection{Monte Carlo Algorithms}
    \paragraph{Metropolis}
        A Metropolis Monte Carlo simulation of an Ising model will
        choose a random spin, calculate the energy change
        \(\Delta \hat H\) defined in eq. \eqref{eq:dH} that would result
        from a flip of and excute the flip with the probability \(A\)
        eq. \eqref{eq:metropolis} \cite{NewmanBarkema1999} \cite{Katzgraber2011}.
        \begin{equation}
            \Delta \hat H = \hat H(\nu) - \hat H(\mu)\\
            \label{eq:dH}
        \end{equation}
        \begin{equation}
            A(\mu \to \nu)
            \begin{cases}
                1                                 & \Delta \hat H \le 0 \\
                \exp{\brac{-\beta \Delta \hat H}} & \Delta \hat H > 0
            \end{cases}
            \label{eq:metropolis}
        \end{equation}
        So if a transition lowers the energy it is always done. This
        results in a good ratio between calculations and transitions.

    \paragraph{Wolff}
        Close to the critical temperature \(T_c\) the Metropolis
        gets slower. This is called \emph{critical slowing} down and the
        cause is beyond the scope of this thesis.\\
        Using a cluster algorithm like the Wolff
        algorithm \cite{Wolff1989} speeds things up.
        For an Ising model the Wolff algorithm builds a cluster of sites
        starting with a random site and adding neighboring sites of the
        same spin with probability \(P_{\mathrm{add}}\) from eq. \eqref{eq:wolffAdd}
        \begin{equation}
            P_{\mathrm{add}} = 1-\exp\brac{-2\beta J}
            \label{eq:wolffAdd}
        \end{equation}
        Where \(J\) is the coupling constant (explained in section
        \ref{ssec:isingmodel}). The neighboring sites of the added sites
        are also considered and so forth. When there are no more sites
        to add, the spin of every site in the cluster is flipped
        \cite[S. ??]{NewmanBarkema1999} \cite[S. 151f]{Katzgraber2011}.
        This leads fast to new uncorrelated states at the critical
        temperature because big clusters are flipped. But there are not
        much advantages at high or low temperatures. At low temperatures
        the cluster will consist of almost all sites such that all but
        very few spins will be flipped. At high temperatures the cluster
        will only contain very few sites.
        Both situaitions have no advantage against the Metropolis algorithm.

    \paragraph{Parallel Tempering}
        %~ The main aim is to obtain the critical temperatures
        %~ \(T_c\) for different disorder paramters \(\sigma\).
        %~ Therefore it is necessary to simulate for many temperatures,
        %~ so that \emph{Parallel Tempering}\footnote{Before R. H.
            %~ Swendsen published this paper, a algorithm \(MC^3\) was
            %~ already published with the same idea. [citation needed]}
        %~ \cite{ParallelTempering1986} is a suited algorithm.
        Parallel Tempering simulates many identical systems at different
        temperatures and periodically swaps the spin configurations
        between two neighboring temperatures with probability \(P\) from
        eq. \eqref{eq:partemp} \cite[S. ??]{NewmanBarkema1999} \cite[S. 155ff]{Katzgraber2011}.
        \begin{equation}
            P((E_i,T_i) \to (E_{i+1},T_{i+1})) = \min\brac{1,\exp\brac{\brac{E_{i+1}-E_i}\brac{\frac{1}{T_{i+1}}-\frac{1}{T_i}}}}
            \label{eq:partemp}
        \end{equation}
        This has the advantages that correlation times of single
        temperatures are far smaller because their spin configuration
        gets often replaced by an other uncorrelated configuration. In
        many cases the more important advantage is, that a system which
        is trapped in a local minimum at a given temperature, can travel
        to higher temperatures, leave it's local minimum and cool down
        again in a lower minimum.

    \paragraph{Implementation Details}
        Here a mixture of the above three algorithms is used.
        Each sweep \(N\) Metropolis spin flips, one Wolff cluster flip
        and one Parallel Tempering swap are performed. Where \(N\) is the
        count of sites.
%~ Monte Carlo Simulationen:\\
    %~ Wolff-Cluster Algorithmus \cite{Wolff1989} (siehe auch \cite[S. xx]{NewmanBarkema1999}),
    %~ Metropolis Sweep \cite{Metropolis1953} (siehe auch \cite[S. xx]{NewmanBarkema1999}),
    %~ Parallel Tempering \cite{ParallelTempering1986} (siehe auch \cite[S. xx]{NewmanBarkema1999} \cite[S. xx]{Katzgraber2011})\\

>>>>

\subsection{Monte Carlo Algorithms}
    \paragraph{Metropolis}
        A Metropolis Monte Carlo simulation of an Ising model will
        choose a random spin, calculate the energy change
        \(\Delta \hat H\) defined in eq. \eqref{eq:dH} that would result
        from a flip of that spin and excute the flip with the probability \(A\)
        eq. \eqref{eq:metropolis} \cite{NewmanBarkema1999} \cite{Katzgraber2011}.
        \begin{equation}
            \Delta \hat H = \hat H(\nu) - \hat H(\mu)\\
            \label{eq:dH}
        \end{equation}
        \begin{equation}
            A(\mu \to \nu)
            \begin{cases}
                1                                 & \Delta \hat H \le 0 \\
                \exp{\brac{-\beta \Delta \hat H}} & \Delta \hat H > 0
            \end{cases}
            \label{eq:metropolis}
        \end{equation}
        So if a transition lowers the energy it is always done. This
        results in a high ratio between chosen spins and flipped spins.
        Therefore it minimizes the calculations needed for a change of
        the state.

    \paragraph{Wolff}
        Close to the critical temperature \(T_c\) the Metropolis
        gets slower. This is called \emph{critical slowing} down and the
        cause is beyond the scope of this thesis.\\
        Using a cluster algorithm like the Wolff
        algorithm \cite{Wolff1989} speeds things up.
        For an Ising model the Wolff algorithm builds a cluster of sites
        starting with a random site and adding neighboring sites of the
        same spin with probability \(P_{\mathrm{add}}\) from eq. \eqref{eq:wolffAdd}
        \begin{equation}
            P_{\mathrm{add}} = 1-\exp\brac{-2\beta J}
            \label{eq:wolffAdd}
        \end{equation}
        Where \(J\) is the coupling constant (explained in section
        \ref{ssec:isingmodel}). The neighboring sites of the added sites
        are also considered and so forth. When there are no more sites
        to add, the spin of every site in the cluster is flipped
        \cite[S. ??]{NewmanBarkema1999} \cite[S. 151f]{Katzgraber2011}.
        This leads fast to new uncorrelated states at the critical
        temperature because big clusters are flipped. But there are not
        much advantages at high or low temperatures. At low temperatures
        the cluster will consist of almost all sites such that all but
        very few spins will be flipped. At high temperatures the cluster
        will only contain very few sites.
        Both situaitions have no advantage against the Metropolis algorithm.\\
        So one would activate this algorithm near the critical temperature
        but use a simple Metropolis algorithm at high and low temperatures.

    \paragraph{Parallel Tempering}
        %~ The main aim is to obtain the critical temperatures
        %~ \(T_c\) for different disorder paramters \(\sigma\).
        %~ Therefore it is necessary to simulate for many temperatures,
        %~ so that \emph{Parallel Tempering}\footnote{Before R. H.
            %~ Swendsen published this paper, a algorithm \(MC^3\) was
            %~ already published with the same idea. [citation needed]}
        %~ \cite{ParallelTempering1986} is a suited algorithm.
        Parallel Tempering simulates many identical systems at different
        temperatures and periodically swaps the spin configurations
        between two neighboring temperatures with probability \(P\) from
        eq. \eqref{eq:partemp} \cite[S. ??]{NewmanBarkema1999} \cite[S. 155ff]{Katzgraber2011}.
        \begin{equation}
            P((E_i,T_i) \to (E_{i+1},T_{i+1})) = \min\brac{1,\exp\brac{\brac{E_{i+1}-E_i}\brac{\frac{1}{T_{i+1}}-\frac{1}{T_i}}}}
            \label{eq:partemp}
        \end{equation}
        This has the advantages that correlation times of single
        temperatures are far smaller because their spin configuration
        gets often replaced by an other uncorrelated configuration. In
        many cases the more important advantage is, that a system which
        is trapped in a local minimum at a given temperature, can travel
        to higher temperatures, leave it's local minimum and cool down
        again in a lower minimum.\\
        In the case of a ferromagnetic Ising model the risk to get trapped
        in an local energy minimum is very low. Though the autocorrelation
        decreases significantly. In the scope of this thesis this comes
        for no cost, because one has to simulate for many temperatures
        to determine the critical temperature. The additional calculations
        to determine wheather to swap configurations or not are small in
        comparison with those that would be needed to generate a new
        uncorrelated state without \emph{parallel tempering}.

    \paragraph{Implementation Details}
        Here a mixture of the above three algorithms is used.
        Each sweep \(N\) Metropolis spin flips, one Wolff cluster flip
        and one parallel tempering swap are performed. Where \(N\) is the
        count of sites.\\
        Because it is not known before, where the critical temperatures
        \(T_c\) are located, the Wolff cluster algorithm is used for
        every temperature. The speed up at criticality is worth the
        moderate slow down at other temperatures.
